---
title: "TCH & Cover~10~"
output: 
      html_document:
          toc: true
          toc_depth: 2
          toc_float: true
          code_folding: hide
---


```{r echo=FALSE}
knitr::opts_chunk$set(cache = T)
```

<style>
body {
text-align: justify}
</style>

<br>

# Lidar data
We have lidar data from Paisagens Sustentáveis (
<a href="https://www.paisagenslidar.cnptia.embrapa.br/webgis/" target="_blank">PaiSus</a>
), Marcus Vinicio Neves d'Oliveira (**MarVin**), Estimativa de biomassa na Amazônia (
<a href="http://www.ccst.inpe.br/projetos/eba-estimativa-de-biomassa-na-amazonia/" target="_blank">EBA</a>
).

From Sassan Saatchi we have 1-m resolution CHMs for lidar flights in Chocó, Colombia.

Canopy Height Models (CHM) were processed only from lidar having field inventories in their flight path.

***TODO:*** *Add map and/or list.*

<br>

# Canopy Height Models

LAStools were used to compute Canopy Height Models (CHM). First, ground points were classified, then height was normalised and to obtain a CHM of 1-m resolution, a pit-free algorithm (based on 
<a href="https://doi.org/10.14358/PERS.80.9.863" target="_blank">Khosravipour <i>et al</i>., 2014</a>
) was applied.

Scripts and pipelines were written in R, using the function `system()` to call LAStools. Since LAStools are Windows binaries, I used 
<a href="https://www.winehq.org/" target="_blank">Wine software</a> 
to run them under macOS Mojave as this is the last 32-bit-compatible operating system 

<br>

## Ground & Height

I followed this blog: 
<a href="https://rapidlasso.com/2014/11/04/rasterizing-perfect-canopy-height-models-from-lidar/" target="_blank">Rasterizing Perfect Canopy Height Models from LiDAR</a> 

Ground points were classified using `lasground` and then height was normalised using `lasheight`, dropping z values above 80m as we know trees in these areas won't be that tall. We can then filter out *e.g.* points sampling birds flying.

```{r }
las.input <- "raw/ANA_A01_2017_laz/*.la{z,s}"
hnorm.dir <- "hnorm/ANA_A01_2017_laz"

callground1 <- paste(
  'wine', '~/LAStools/bin/lasground_new',
  '-i', las.input,
  '-wilderness',
  '-odix _ground',
  '-olaz -odir', hnorm.dir
  )
```

We paste all arguments and flags into a text string and then run it through the ```system()``` function:

```{r, eval=FALSE}
system(callground1)
```

It will eventually call this command into the shell Terminal: 

````
`r paste('$',callground1)`
````

<br>

## CHM computation

To create CHMs, a grid is superposed on top of the point cloud to leave the highest first-return per grid cell. Because some grids (pixels) may remain empty (pits), points are converted to disks of a certain diameter to give higher chances to intersect with a pixel. A Triangulated Irregular Network (TIN) is generated with those and then this TIN is converted to a raster. The pit-free algorithm based on 
<a href="https://doi.org/10.14358/PERS.80.9.863" target="_blank">Khosravipour <i>et al</i>., 2014</a>)
creates a sequence of partial CHMs with finer resolution under different height thresholds. Then, these are stacked together leaving the highest pixel values.

<br>

Some parameters were specified:


| | |
|:-:|:-|
|**resol**|Final resolution of the CHM|
|**substep**|Resolution of grid in the steps previous to the final CHM|
|**subcircle**|Diameter splatting points into disks|
|**kill**|Maximum TIN triangle size allowed when creating the pit-free partial CHMs|
|**h_seq**|Height thresholds for the pit-free sequence|
|

***TO-DO: clean this script...***

```{r , eval=FALSE}
las.input <- "raw/ANA_A01_2017_laz/*.la{z,s}"
hnorm.dir <- "hnorm/ANA_A01_2017_laz"

### compute CHM ----
resol = 1
subcircle = 0.2
substep = 0.2
kill = 2.0
h_seq = c(2,5,10,20,30,40,50,60)

### thin las -----
# use multicore, set to the number of our cores
numCores <- detectCores()
cl <- makeCluster(numCores)
registerDoParallel(cl)

foreach(i = 1:length(base_folders)) %dopar% {
  ### create folders -----
  out.folder <- 
    paste0('~/Rproj/07_carbon_br/data/lidar/',
           'chm_pitfree_subcir',sub("\\.", "", as.character(subcircle)),
           '_kill',kill,
           '_substep',sub("\\.", "", as.character(substep)))
  dir.create(out.folder ,recursive = T)
  (temp_laz.dir <- paste0(out.folder,'/temp_laz/PaiSus/',base_folders[i]))
  dir.create(temp_laz.dir ,recursive = T)
  ### thin las -----
  thin_las <- 
    paste(
      'wine ~/LAStools/bin/lasthin', 
      '-i',paste0(hnorm.dir[i],'/*hnorm.la{z,s}'),
      '-subcircle', subcircle, 
      '-step', substep, 
      '-highest -olaz',
      '-odir', temp_laz.dir
    )
  cat('--Calling: \n',thin_las,'\n\n')
  system(thin_las)
}
stopImplicitCluster()
### chm ground and interpolation -----
numCores <- detectCores()
print(cl <- makeCluster(numCores))
registerDoParallel(cl)

foreach(i = 1:length(hnorm.dir)) %dopar% {
  # create folders -----
  out.folder <- 
    paste0('~/Rproj/07_carbon_br/data/lidar/',
           'chm_pitfree_subcir',sub("\\.", "", as.character(subcircle)),
           '_kill',kill,
           '_substep',sub("\\.", "", as.character(substep)))
  dir.create(out.folder ,recursive = T)
  (temp_laz.dir <- paste0(out.folder,'/temp_laz/PaiSus/',base_folders[i]))
  dir.create(temp_laz.dir ,recursive = T)
  (tmp_chm.dir <- paste0(out.folder,'/tmp_chm/PaiSus/',base_folders[i]))
  dir.create(tmp_chm.dir ,recursive = T)
  (chm_pitfree.dir <- paste0(out.folder,'/chm_pitfree/PaiSus/'))
  dir.create(chm_pitfree.dir ,recursive = T)
  # chm ground ----
  chm_ground <- 
    paste(
      'wine ~/LAStools/bin/blast2dem',
      '-i',paste0(hnorm.dir[i],'/*hnorm.la{z,s}'),
      '-merged -drop_z_above 0.1',
      '-step', resol,
      '-o', paste0(tmp_chm.dir,'/chm_ground.bil')
    )
  cat('--Calling: \n',chm_ground,'\n\n')
  system(chm_ground)

  # interpolation ----
  chm_interp0 <- 
    paste(
      'wine ~/LAStools/bin/blast2dem',
      '-i', paste0(temp_laz.dir,'/*.la{z,s}'),
      '-merged -v',
      '-step', resol,
      '-kill', kill, 
      '-o', paste0(tmp_chm.dir,'/chm_00.bil')
    )
  cat('--Calling: \n',chm_interp0,'\n\n')
  system(chm_interp0)
  
  chm_interpol <- 
    paste(
      paste(
        'wine ~/LAStools/bin/blast2dem',
        '-i', paste0(temp_laz.dir,'/*.la{z,s}'),
        '-merged -v',
        '-step', resol,
        '-kill', kill),
      paste('-drop_z_below',
            h_seq, '-o', paste0(tmp_chm.dir,'/chm_',sprintf('%02d',h_seq),'.bil'))
    )
  cat('--Calling: \n',paste(chm_interpol,collapse = '\n'),'\n\n')
  system(paste(chm_interpol,collapse = '\n'))
  
  
  # merge chms png and asc ----
  merge_chms <- 
    paste(
      'wine ~/LAStools/bin/lasgrid',
      '-i', paste0(tmp_chm.dir,'/*.bil'),
      '-merged -highest -false -v', 
      '-step', resol,
      '-o', paste0(chm_pitfree.dir,'/',basename(tmp_chm.dir),'__',
                   sub("\\.", "", as.character(subcircle)),kill,
                   sub("\\.", "", as.character(substep)),'.png')
    )
  cat('--Calling: \n',merge_chms,'\n\n')
  system(merge_chms)
  
  merge_chms_asc <- 
    paste(
      'wine ~/LAStools/bin/lasgrid',
      '-i', paste0(tmp_chm.dir,'/*.bil'),
      '-merged -highest -v', 
      '-step', resol,
      '-o', paste0(chm_pitfree.dir,'/',basename(tmp_chm.dir),'__',
                   sub("\\.", "", as.character(subcircle)),kill,
                   sub("\\.", "", as.character(substep)),'.asc')
    )
  cat('--Calling: \n',merge_chms_asc,'\n\n')
  system(merge_chms_asc)
}
stopImplicitCluster()

```

<br>

# Top-of-Canopy Height (TCH)

After creating CHMs for the lidar scenes, Top-of-Canopy Height (TCH) was extracted for each forest inventory polygon. Basically, selecting the pixels within the polygon and then calculating the mean value of those.

<br>

## TCH plots

Histogram?

<br>

# Cover~10~

A cover fraction was acquired by calculating the proportion of pixel values ≥ 10 m within a plot.

<br>

# CHM by plot{.tabset}

***TODO:*** *change to drop-down menu & add pdf link*

You can see the CHM image(s) in each plot. There are some plots that have several lidar resurveys.


```{r , echo=FALSE, message=FALSE, warning=FALSE, results='hide', cache=TRUE} 
library(raster)
library(sf)
fr <- list.files('~/Rproj/07_carbon_br/data/lidar/chm_pitfree_subcir02_kill2_substep02/chm_pitfree/', recursive = T, pattern = 'asc', full.names = T)
nr <- sub('.asc', '', basename(fr))
# rl <- lapply(fr, function(x) raster(x))
rr <- readRDS("~/Rproj/07_carbon_br/data/lidar/chm_pitfree_subcir02_kill2_substep02/tmpRmd/rxmk_v1.Rds")
# nr <-  sapply(rl, names)

dbptch <- st_read('~/Rproj/07_carbon_br/data/inventories/plots_clean/shapes/db_parcela_all_tch/dbptch_v1/dbptch_v1.shp')
names(dbptch) <- read.csv('~/Rproj/07_carbon_br/data/inventories/plots_clean/shapes/db_parcela_all_tch/dbptch_v1/dbptch_v1_varnames.csv')[,1]

tmp <- gregexpr('_',dbptch$chmintersect)
nchm <- substr(dbptch$chmintersect,1,sapply(tmp,'[',3)-1)
```


```{r, 'chms_plots', results='asis', echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE} 
cat('\nSelect the project \n\n')
for(k in unique(dbptch$owner)){
  cat('##',k,'{.tabset}\n')
  cat('\nSelect the location \n\n')
  lcid <- unique(dbptch$location[dbptch$owner==k])
  
  for(j in lcid) {
    cat('###',j,'{.tabset .unnumbered}\n')
    cat('\nSelect the plot \n\n')
    plid <- dbptch$plotID[dbptch$location==j]
    
    for(i in unique(plid)) {
      cat('####',i,'{.tabset .unnumbered}\n')
      cat('\nSelect the CHM \n\n')
      chmid <- which(paste(dbptch$owner,dbptch$location,dbptch$plotID)%in%paste(k,j,i))
      
      for(h in chmid) {
        cat('#####',nchm[h],'{.unnumbered}\n')
      
        px <- dbptch[h,]
        rx <- rr[[h]]
        # rx <- rl[[grep(dbptch$chmintersect[h], nr)]]
        # rxc <- mask(rl[[grep(dbptch$chmintersect[h], nr)]], dbptch[h,])
        # rxc <- trim(mask(rx,px))
        # print(rx)

        plot(st_geometry(px), col=NA, border=NA, main=names(rx))
        plot(rx, zlim=c(0,78.4), col=lidR::height.colors(100), add=T)
        plot(st_geometry(px), add=T)
        text(st_coordinates(st_centroid(px)),adj=.5,paste0('Lidar:',
                                                          '\nTCH= ',round(px$TCHmean,1),
                                                          '\nCover10= ',round(px$cover10,2),
                                                          '\nN.pixel= ',px$npx,
                                                          '\narea.shp= ',round(st_area(px)/10000,3),' ha',
                                                          '\n\nField:\n',px$location,' ',px$year," (",px$plotID,")",
                                                          '\nBA= ',round(px$BAplot,1),
                                                          '\nWD= ',round(px$WDplot2,1),
                                                          '\nACD= ',round(px$ACD,1)
                                                          ))

      cat(' \n \n')
    }
      cat(' \n \n')
    }
    cat(' \n \n')
  }
  cat(' \n \n')
}

```


