---
title: "Carbon project"
date: "Last update 11/05/2021"
output: 
      html_document:
          toc: true
          toc_float: true
---

<style>
body {
text-align: justify}
</style>

Welcome, the purpose of this website is to keep you updated!

While other datasets are yet to come, I am working with forest inventory data plus matching lidar from **Paisagens Sustentáveis** and **Marcus Vinicio Neves d'Oliveira** in Brazil. 


## Cleaning data
Inventory datasets were compiled and standardised. Taxonomic errors were corrected using the Taxonomic Name Resolution Service v5.0 (https://tnrs.biendata.org/). 

All inventories matching with lidar (`````r nrow(cars)`````plots/tansects in `````r nrow(cars)````` locations) were selected, but some data were excluded sequentially:

- ```r nrow(cars)``` inventory transects (*i.e.*, 500 meter transects in Paisagen Sustentáveis).

- ```r nrow(cars)``` re-sampled inventory plots. I selected most recent plots (not always) with the least time difference from its lidar survey.

- ```r nrow(cars)``` plots which had more than half of the trees dead and less than 10 individuals.

- ```r nrow(cars)``` dead individuals were removed.  *(I may add them later into biomass estimations, ```r nrow(cars)``` out of ```r nrow(cars)``` (```r nrow(cars)```%) dead trees have DBH measurements)*.

- ```r nrow(cars)``` lianas.

- ```r nrow(cars)``` trees without DBH measurement.


Finally, what's left for the analysis are ```r nrow(cars)``` individuals in ```r nrow(cars)``` plots in ```r nrow(cars)``` locations:

```{r}
#Add a table or something... with plot size...
summary(cars$dist)
summary(cars$speed)
```

## Map of inventory and lidar boundaries
The following map shows plot locations and lidar transects available for the analysis:
```{=html}
<iframe src="https://www.google.com/maps/d/u/0/embed?mid=1B8fDAx8gQiWn6hD3BzQkBcjWoqBajU1E" width="640" height="480"></iframe>
```



# AGB estimations

In order to calculate Above-ground biomass (AGB) from field inventories, allomtric ecuations are used. In this case, Chave et al. (201X) equation 4 was used:
$$
AGB = DBH^2...
$$
Further cleaning of data...


## DBH (Diameter at Breast Height)

Plots from Paisagens Sustentáveis have different DBH threshold in their plots. Many of them, measured trees >35cm DBH in the whole plot area and >10cm in a sub-plot area. Sub-plots vary in size across locations, for instance, a 10m x 50m subplot within a 50m x 50m plot. Therefore, some expansion factors will be applied when calculating plot AGB.

Add image? Yes...

```{r}
#Add table of min DBH in plots (excel)
summary(cars$dist)
summary(cars$speed)
```


Minimum DBH threshold is consisteng in Paisagens Sustentáveis where few DBHs are below 10 cm, therefore I considered them typing errors. I thought of giving them the mean DBH value for the species but I ended up excluding them for the moment.

Data in MarVin is not fully consistent and minimum DBH threshold seem to go from 5cm, but not all plots. *I need to get proper metadata from him.* Also, there are few points below 5cm which I consider they are typos.

```{r}
#Add a table with DBH cut() and owner
summary(cars$dist)
summary(cars$speed)
```

Now, to make projects comparable for the moment, all trees DBH < 10cm were excluded (```r nrow(cars)``` ind), but this will probably change to a minimum of 5cm DBH in the future, using expansion factors for AGB in plots with no individuals below 10cm. This because it will be a problem *e.g.* when comparing our data with Asner & Mascaro (201X)'s which go from 5cm.

This is the distribution of DBH in our data (*i.e.*, DBH<10cm):
```{r}
#Add histogramof DBH
summary(cars$dist)
summary(cars$speed)
```


## Height

Height values are few among inventories. Only PaiSus had some height values, whereas MarVin data had none.

```{r}
# add number of height values in data, etc
summary(cars$dist)
summary(cars$speed)
```

Therefore, I estimated height of the trees from DBH measurements by constructing local height models.


### Local Height Models

The relationship between DBH and height are different for trees and palms, so I each group had its own local height model. There were ~30 crazy outliers which I did not account for when building the models. *e.g.*, $height = 48.2m$ when $DBH = 12.7 cm$.

```{r}
#Add plot H ~ DBH for trees & palms. show outliers?
summary(cars$dist)
summary(cars$speed)
```


#### Trees





#### Palms





## Wood Density
Wood Density (rho) was acquired from database XX...

In case the species was not found, the mean WD of the genus was allocated. If tree not identified to genus, the mean of the WD of the plot was selected.

```{r}
# Stats ho many were to species, genus, plot
summary(cars$dist)
summary(cars$speed)
```


## AGB calculation
The formula was....



## Error propagation
Markov chains...





# TCH lidar
...





# Basal Area & WD_plot
...




# Modelling
...






# Project description

whose model parameters a and b vary with forest types. Basal Area (BA) and wood density (ρ) will be coupled as sub-models to develop a generalised model that can be applicable across different forest types:

$$
ACD_{plot} = a · TCH^d · BA^c · ρ^b; 
$$
ACD will be calculated using field plot data by summing above-ground biomass (AGB) for each tree (Chave et al., 2014) and then multiplying by 0.47, which is the conversion factor for carbon content (Martin & Thomas, 2011). 

LiDAR data will be used to calculate TCH, while BA and ρ are derived from field plot data. This modelling approach will provide a better understanding on how BA and ρ affect ACD.

