---
title: "Estimating carbon from lidar"
date: "Last update `r format(Sys.time(), '%d-%b-%Y')`"
output: 
      html_document:
          toc: true
          toc_float: true
---


```{r echo=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

<style>
body {
text-align: justify}
</style>

<br>



Hi! I created this website to keep you updated with the project! First, I give an overview of what we want to develop and then the different processes and analyses I have done so far. Main tabs of this website explain further how I processed data for each of the elements of the model (*i.e.*, AGB, TCH, BA & ρ) as well as the modelling process.

<br>

# Project overview

The main goal is to generate equations to estimate aboveground biomass/carbon (AGB/ACD) from lidar. We want to better understand variation in Wood Density (ρ) and Basal Area (BA) to improve the Top-of-the-Canopy Height (TCH) - Biomass relationships. The aim is to generate models based on the following expression:


$$
ACD_{plot} = a · TCH^b · BA^c · ρ^d
$$

ACD will be calculated using field plot data by summing [above-ground biomass](01_agb.html) (AGB) for each tree (Chave et al., 2014) and then multiplying by 0.48, which is the conversion factor for carbon content (Martin & Thomas, 2011). 

LiDAR data will be used to [calculate TCH](02_tch.html), while [BA](03_ba.html#wood-density-(plot)) and [ρ (*aka* WD)](03_ba.html#basal-area-(plot)) are derived from field plot data. 
<!-- Our aim is to provide a better understanding on how BA and ρ affect ACD-TCH relationships to improve carbon estimations. -->

<br>

# Data

## Data collection

While other datasets are yet to come, I am working with forest inventories plus matching lidar from Paisagens Sustentáveis (**PaiSus**) and Marcus Vinicio Neves d'Oliveira (**MarVin**) in Brazil. 

Other colaborators will hopefully provide their datasets.

*NOTE: Sassan Saatchi shared data for the Colombian Chocó but I still have to incorpore those into the analysis*


<br>

## Cleaning inventory data

```{r echo=F, results='hide'}
INV <- read.csv('~/Rproj/07_carbon_br/data/inventories/plots_clean/all_inventories_v1_resoltaxo.csv', stringsAsFactors = F)
np0 <- length(table(INV$parcela))
nl0 <- length(table(INV$location))

ntranp <- table(unique(INV[grep('500',INV$sizeplot), c('parcela','sizeplot')])$sizeplot)
INV <- INV[!grepl('500',INV$sizeplot),]

## db parcela (dbp) creation and selecting the plots for the analysis
dbp <- unique(INV[c("state","Name","location","plotID","sizeplot","parcela","year","owner")])
(npini <- nrow(dbp))

cndbp <-  !(dbp$location=='ANA_A01' & dbp$year==2018) & 
          !(dbp$location=='CAU_A01' & dbp$year==2014) & 
          !(dbp$location=='CAU_A01' & dbp$plotID%in%c('T01_P3','T01_P4','T05_P1')) & 
          !(dbp$location=='JAM_A02' & grepl('T03|T01_P1',dbp$plotID)) & 
          !(dbp$location=='PAR_A01' & dbp$plotID=='P13') & 
          !(dbp$location=='SAN_A01' & dbp$year==2016) & 
          !(dbp$location=='SAN_A02' & dbp$plotID=='G01_P04') & 
          !dbp$location=='TAC_A01' & 
          !(dbp$location=='TAP_A01' & dbp$year%in%c(2018,2016)) & 
          !(dbp$location=='TAP_A03' & dbp$plotID=='P59')
631-32-88-3-5-1-8-1-27-11-1
dbp <- dbp[cndbp,]
npfin <- nrow(dbp) # Se quitaron 177 parcelas
npini-npfin

INV <- INV[INV$parcela%in%dbp$parcela,]

# Number of individuals
z <- data.frame(table(INV$parcela))
(names(z) <- c('parcela','n'))
dbp <- merge(dbp, z, by = 'parcela')

# Number of Dead
z <- table(INV$parcela,INV$dead)
dbp <- merge(dbp, data.frame(alive=z[,1], dead=z[,2], parcela=row.names(z)), by = 'parcela')
dbp$dead.percent <- round(dbp$dead/dbp$n*100,1)

table(INV$dead, useNA = 'ifany')
table(!is.na(INV$DBH[INV$dead])) # there are many dead individuals with DBH

npdead <- nrow(dbp[dbp$dead.percent >= 50 & dbp$n <= 10,])
dbp <- dbp[!(dbp$dead.percent >= 50 & dbp$n <= 10),] # 8 plots less. Aca se perdieron unas parcelas porque todos sus individuos estan muertos (ie, algunas de FN)
INV <- INV[INV$parcela%in%dbp$parcela,]

niddbh <- table(!is.na(INV$DBH[INV$dead])) # there are many dead individuals with DBH
INV <- INV[!INV$dead,] 

length(dbp$parcela)==length(unique(INV$parcela))
table(dbp$parcela%in%unique(INV$parcela))

table(INV$tree.type, useNA = 'ifany')
niL <- nrow(INV[INV$tree.type=='L',])
INV <- INV[INV$tree.type!='L',]

##Number of DBH=NA
table(is.na(INV$DBH))
ninadbh <- nrow(INV[is.na(INV$DBH),])
INV <- INV[!is.na(INV$DBH),]


```

Inventory datasets were compiled and standardised from collaborators shown above. Taxonomic errors were corrected using the Taxonomic Name Resolution Service v5.0 (https://tnrs.biendata.org/). 

All inventories matching with lidar (```r np0```plots/tansects in ```r nl0``` locations) were selected, but some data were excluded in the following sequence:

- ```r sum(ntranp)``` inventory transects (*i.e.*, 500 meter transects in Paisagen Sustentáveis).

- ```r npini-npfin``` re-sampled plots. I selected most recent (not always) with the least time difference from its lidar survey.

- ```r npdead``` plots which had more than half of the trees dead and less than 10 individuals.

- ```r sum(niddbh)``` dead individuals were removed.  *(I may add them later into biomass estimations, ```r niddbh[2]``` dead trees (```r round(niddbh[2]/sum(niddbh)*100,0)```%) have DBH measurements)*.  
  *NOTE: MarVin dead trees are...*

- ```r niL``` lianas.

- ```r ninadbh``` trees without DBH measurement.

<br>

Finally, what's left for the analysis are ```r nrow(INV)``` individuals in ```r length(unique(INV$parcela))``` plots in ```r length(unique(INV$location))``` locations:

```{r echo=F}
#Add a table or something... with plot size...
message('Number of individuals by size and project:')
table(INV$owner,INV$sizeplot)

message('Number of plots by size and project:')
table(dbp$owner,dbp$sizeplot)

message('Number of plots by size and project-location:')
table(paste(dbp$owner,dbp$location),dbp$sizeplot)

```

<br>

## Map of inventory and lidar boundaries

The following map shows plot locations and lidar transects available for the analysis: *(NOTE: this is an outdated map. MarVin and Sassans plots are not included)*



```{=html}
<iframe src="https://www.google.com/maps/d/u/0/embed?mid=1B8fDAx8gQiWn6hD3BzQkBcjWoqBajU1E" width="640" height="480"></iframe>
```








